---
title: "Add Pset name here"
author: "Your Name Here"
date: "Date Here"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

<!-- .Rmd files use  markdown, a text mark up language, to provide formating.-->

<!--Text include within these strange arrows are comments and will not show up when you knit-->

# Front matter

This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: \<\<\_\>\>

Upload your collaborators here: https://forms.gle/3mH1ofZZj9ZCgBHb8

Late coins used this pset: 0. Late coins left: 4. <!--You may use one for a given assignment.-->

# R for Data Science Exercises

<!--Note: Please do not restate questions in your submissions, directly write your answer--->

## First Section

<!--(Notice the use of two `##`)-->

1.  YOUR TEXT ANSWER GOES HERE (Please delete)

-   The dash symbol creates unordered list

-   Another example

    1.  numbering creates ordered list
    2.  Another example

    -   like This
    -   Like this as well

**Delete all the text in `1.` before you write your answer**

```{r}
# YOUR CODE GOES HERE  (Please delete)
ggplot(mpg, aes(manufacturer)) + geom_bar()
```


## Second Section

1.  YOUR TEXT ANSWER GOES HERE (Please delete)

```{r}
# YOUR CODE GOES HERE  (Please delete)
ggplot(mtcars, aes(factor(cyl), mpg)) + geom_violin(aes(fill=mpg))
```

