---
title: "applied_PS4"
author: "Zara & Matthew"
date: "13th May 2023"
output: pdf_document
---
<<<<<<< HEAD
=======
<!-- .Rmd files use  markdown, a text mark up language, to provide formating.-->

<!--Text include within these strange arrows are comments and will not show up when you knit-->


# Front matter

This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: M.S & Z.N

Upload your collaborators here: https://forms.gle/3mH1ofZZj9ZCgBHb8

Late coins used this pset: 0. Late coins left: 4. <!--You may use one for a given assignment.-->


```{r message=FALSE, warning=FALSE}
rm(list=ls())
install.packages("anytime")
install.packages("bigrquery")
install.packages("stringr")
install.packages("ggmap")
library(bigrquery)
library(tidyverse)
library(testthat)
library(ggplot2)
library(dplyr)
library(lubridate)
rm(list=ls())
install.packages("readr")

```


## 1 Prelim Question 
1.1 Yes, waze data set has been deleted from my computer 

1.2  One additional variable that Waze could track, which is not available 
currently, is "Weather Conditions." This would provide real-time information on 
weather conditions along routes, enhancing user safety, improving route 
planning, and enhancing the overall user experience. Waze could partner with 
weather data providers, enable user reporting, or integrate weather sensors to 
gather accurate and timely weather information.

The variable "nThumbsUp" could potentially be tracked and utilized better by
Waze. Currently, Waze allows users to give thumbs-up ratings to reports, but the 
specific count of thumbs-up is not always visible or utilized prominently 
in the application. Enhancing the tracking and visibility of the "nThumbsUp" 
variable would provide valuable information about the credibility and 
reliability of user-generated reports. It could help prioritize and validate 
reports based on user feedback, ensuring more accurate and reliable information
for other Waze users.

1.3 
Self-selection among Waze users influences the data present on the platform.
Users who actively engage with the app are more likely to contribute 
traffic-related information, leading to a higher density of such data. However, 
less popular features and rural areas may have limited data due to user 
preferences and geographic concentration. Understanding these biases is crucial 
for using Waze data effectively.

## 2 git practice

# 1

The merge conflict occurred because both Partner A and Partner B made changes 
to the same line of the file in their respective branches, which conflicted 
with each other. As a result, when Partner A tried to merge Partner B's 
changes into their branch, there was a conflict because the same line was 
changed in two different ways. This requires manual intervention to resolve 
the conflict by editing the file to incorporate both changes.

## 3 Obtaining data from cloud using SQL

# 1

```{r message=FALSE, warning=FALSE}

install.packages("readr")
library(readr)
dfw <- read_rds("/Users/mattida6/Desktop/GitHub/waze_data.rds")

```

## 4 Data exploration in JSON

# 1

It appears that the data format being used is JSON (JavaScript Object Notation). 
JSON is a text-based data interchange format that is used to represent data in 
a structured manner. JSON is different from other file formats such as 
CSV (Comma-Separated Values) in several ways:

1. Structure: CSV is a flat file format that uses a comma (or other delimiter) 
to separate values in a row, while JSON is a hierarchical data format that uses 
nested objects and arrays to represent complex data structures.

2. Data Types: CSV typically represents all data as text, while JSON supports 
a wider range of data types, including numbers, booleans, strings, arrays, 
and objects.

3. Readability: JSON is often easier to read and understand than CSV 
because it uses a structured format with clearly defined objects and attributes.

4. Flexibility: JSON is a more flexible format than CSV because it can 
represent complex data structures and allows for the inclusion of metadata,
such as attribute names and data types.

Overall, JSON is a more robust and versatile data format than CSV,
making it well-suited for applications that require complex data structures 
and the ability to include metadata. However, CSV remains a popular format for 
simple data sets because of its simplicity and wide support.

# 2

```{r message=FALSE, warning=FALSE}
library(jsonlite)
lf_json <- fromJSON("/Users/mattida6/Desktop/stats_code/69c3e9b3-182c-4ec9-a5f3-c0e176568a3d.json")
```

As for why we can't convert the JSON feed directly to a tibble, it's because 
a tibble requires a rectangular data structure, where each column has the same
length and data type. In contrast, JSON can represent complex hierarchical data 
structures with varying data types. While it is possible to convert a JSON 
object to a tibble using the as_tibble() function from the tidyjson package, 
this requires some additional data wrangling to ensure that the resulting 
tibble is rectangular and well-structured.

# 3

```{r message=FALSE, warning=FALSE}
# Get the length of the list
length(lf_json)

# Get the names of each item in the list
names(lf_json)

# Check which items in the list can be converted to tibbles
unique(sapply(lf_json, class))
```

## 5 Data Cleaning 

5.1 Data downloaded from convas in the form of an rds file. computed a test_that 
that the dataset has 737,357 rows test passed. 
```{r}
test_that(
  "we have the right number of row",
  expect_equal(nrow(dfw),737357))

```
5.2 The data covers the city of chicago 
```{r}

city_number<-dfw|>
  distinct(city)|>
  print()
```
5.3 Separate columns for longitude and latitude created. 
```{r}
dfw <- dfw |>
  mutate(longitude = as.numeric(str_extract(geo, "-?\\d+\\.\\d+(?=\\s)")),
         latitude = as.numeric(str_extract(geo, "(?<=\\s)-?\\d+\\.\\d+")))|>
  relocate(longitude, latitude)|>
  print()



```
5.4 Alerts are most common at night post from 20:00 hours to 22:00 hours.

```{r}

library(ggplot2)
library(dplyr)
library(lubridate)

dfw <- dfw |>
  mutate(hour = hour(ymd_hms(ts)))

ggplot(dfw, aes(x = hour, fill = type)) +
  geom_bar() +
  labs(x = "Hour of Day", y = "Number of Alerts", fill = "Alert Type") +
  ggtitle("Number of Alerts by Hour of Day") +
  theme_minimal()+ 
  scale_x_continuous(breaks = seq(0, 23, by = 1))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

5.5 By default the alerts are in central time zones. Column with alerts 
recorded to central time and a second column with alerts on central time 
rounded to thenearest 5 minutes added. 

```{r}
dfw <- dfw |>
  mutate(
    ts = as.POSIXct(ts, format = "%Y-%m-%d %H:%M:%S", tz = "UTC"),
    ts_ct = with_tz(ts, "America/Chicago"),
    ts_ct_rounded = round_date(ts_ct, "5 minutes")
  )


```

## 6 Waze vision zero
## 6 Waze Vision Zero

Partner A 

6.1.1 branch called stony island created. 


6.1.2 

```{r}

library(ggmap)

unique_values <- unique(dfw$street)
unique_values 

corridor_accidents <- dfw[dfw$street == "S Stony Island Ave" & dfw$type == "ACCIDENT", ]

map <- get_stamenmap(
  bbox = c(
    left = min(corridor_accidents$longitude, na.rm = TRUE),
    bottom = min(corridor_accidents$latitude, na.rm = TRUE),
    right = max(corridor_accidents$longitude, na.rm = TRUE),
    top = max(corridor_accidents$latitude, na.rm = TRUE)
  ),
  maptype = "toner-lite",
  zoom = 10
)


ggmap(map) +
  geom_point(data = corridor_accidents, aes(x = longitude, y = latitude), color = "orange", size = 2) +
  labs(title = "Accidents on High Crash Corridor # 25 S Stony Island Ave") +
  theme_minimal()

```

6.1.3 Around this "41.78876_-87.586654" interaction accidents are most common, 
this is near the 59th street, which is not within the boxed parameter 

```{r}


library(ggmap)
library(dplyr)

# Defining the bounding box 
bbox_left <- -87.5876
bbox_bottom <- 41.7799
bbox_right <- -87.5833
bbox_top <- 41.7949


corridor_accidents_bbox <- corridor_accidents |>
  filter(street == "S Stony Island Ave" & 
           longitude >= bbox_left & longitude <= bbox_right & 
           latitude >= bbox_bottom & latitude <= bbox_top)

# Grouping accidents by intersection and counting occurrences
intersection_counts <- count(corridor_accidents_bbox, latitude, longitude)

# Retrieving intersection with the highest number of accidents
most_common_intersection <- with(intersection_counts, paste(
  latitude[which.max(n)], longitude[which.max(n)], sep = " "))

most_common_intersection
```

# 6.2 Partner B

# 6.2.2

W Chicago Ave & N Western Ave, Chicago, IL 60622 = 41.933930, -87.687884
W George St & N Western Ave, Chicago, IL 60618 = 41.895741, -87.686892


```{r message=FALSE, warning=FALSE}
install.packages("ggmap")
library(ggmap)


```

```{r message=FALSE, warning=FALSE}
library(ggmap)
install.packages("ggspatial")
library(ggspatial)

corridor_accidents <- dfw[dfw$street == "N Western Ave" & dfw$type == "ACCIDENT",]
map <- get_stamenmap(bbox = c(left = min(corridor_accidents$longitude),
                              bottom = min(corridor_accidents$latitude),
                              right = max(corridor_accidents$longitude),
                              top = max(corridor_accidents$latitude)),
                     maptype = "toner-lite")

ggmap(map) +
  geom_point(data = corridor_accidents, aes(x = longitude, y = latitude), color = "orange", size = 3) +
  labs(title = "Accidents on High Crash Corridor #: N Western Ave",
       x = "Longitude", y = "Latitude") +
  theme_minimal() +
  annotation_scale(location = "bl", width_hint = 0.3, text_cex = 0.8, bar_step = 0.25) +
  guides(color = FALSE, size = guide_legend(title = "Accidents"))

```

# 6.2.3

No, the coordinates 41°59'26.6"N 87°41'22.8"W are not between
W Chicago Ave & N Western Ave, Chicago, IL 60622 (41.933930, -87.687884) and
W George St & N Western Ave, Chicago, IL 60618 (41.895741, -87.686892). 
These coordinates are actually located much further west, 
near the intersection of N Milwaukee Ave and W Montrose Ave.


# 6.3
# 1
success

## 7 Waze single event
# 1
```{r message=FALSE, warning=FALSE}
dfw |> filter(uuid == "a42bc14b-e080-4621-9221-29dd86e553ce") |> print()
```

A.
```{r message=FALSE, warning=FALSE}
lat_range <- c(41.930, 41.940)
lon_range <- c(-87.695, -87.680)

event_df<- dfw %>%
  filter(
         latitude >= lat_range[1], latitude <= lat_range[2],
         longitude >= lon_range[1], longitude <= lon_range[2],
         type %in% c("ACCIDENT", "JAM")) %>%
  select(uuid, latitude, longitude, type) %>% print()


```

B.
Two men were killed after a wrong-way crash on DuSable Lake Shore Drive Sunday 
morning.

C.
```{r message=FALSE, warning=FALSE, cache=FALSE}
ddfw <- dfw %>%
  mutate(ts = as.POSIXct(ts, format = "%Y-%m-%d %H:%M:%S UTC"), # ensure ts column is in correct format
         date = as.Date(ts), # create new column for date only
         time = format(ts, format = "%H:%M:%S")) %>% # create new column for time only
  select(-ts) # delete the original ts column
head(ddfw)
```

```{r message=FALSE, warning=FALSE}
sixoclock_jams<- ddfw %>%
  filter(
         latitude >= lat_range[1], latitude <= lat_range[2],
         longitude >= lon_range[1], longitude <= lon_range[2],
         type %in% c("JAM")) %>%
  select(uuid, latitude, longitude,type, time) %>% print()
```

```{r message=FALSE, warning=FALSE}
library(lubridate)

ddfw_filtered <- sixoclock_jams %>%
  filter(hour(as.POSIXlt(time, format = "%H:%M:%S")) >= 6,
         hour(as.POSIXlt(time, format = "%H:%M:%S")) < 18) %>% print()
```

```{r message=FALSE, warning=FALSE}
# Create a new date-time column by pasting today's date with the "time" column
ddfw_filtered$datetime <- as.POSIXct(paste(Sys.Date(), ddfw_filtered$time), format="%Y-%m-%d %H:%M:%S")

# Filter the data for the times between 6AM and 6PM
ddfw_filtered <- ddfw_filtered %>%
  filter(hour(datetime) >= 6, hour(datetime) < 18)

# Group the data by time intervals of 1 hour and count the number of occurrences
counts <- aggregate(x = ddfw_filtered["type"], by = list(cut(ddfw_filtered$datetime, breaks = "1 hour")), FUN = length)

# Create a plot of the counts over time
plot(counts$Group.1, counts$type, type = "l", xlab = "Time", ylab = "Count", main = "Number of Jams 6AM-6PM")

```
D.
```{r message=FALSE, warning=FALSE}
# Load the data into a data frame
ddfw_filtered <- data.frame(
  type = c("JAM", "JAM", "JAM", "JAM", "JAM", "JAM"),
  subtype = c("Accident", "Accident", "Heavy traffic", "Road closure", "Accident", "Heavy traffic"),
  time = c("08:00:00", "09:15:00", "10:30:00", "13:00:00", "15:45:00", "17:30:00")
)

# Convert time to POSIXct format
ddfw_filtered$time <- as.POSIXct(ddfw_filtered$time, format = "%H:%M:%S")

# Define weights for each subtype
subtype_weights <- c("Accident" = 3, "Road closure" = 2, "Heavy traffic" = 1)

# Filter data by time between 6am and 6pm
ddfw_filtered <- subset(ddfw_filtered, format(time, "%H") >= "06" & format(time, "%H") < "18")

# Calculate the severity score for each JAM alert
ddfw_filtered$score <- with(ddfw_filtered, subtype_weights[subtype] * table(subtype)[subtype])

# Sum the scores to obtain the total severity score
total_score <- sum(ddfw_filtered$score)

# Print the results
print(ddfw_filtered)
cat("Total severity score:", total_score, "\n")

```
E.
```{r message=FALSE, warning=FALSE}
# Load the data into a data frame
ddfw_filtered <- data.frame(
  type = c("JAM", "JAM", "JAM", "JAM", "JAM", "JAM"),
  subtype = c("Accident", "Accident", "Heavy traffic", "Road closure", "Accident", "Heavy traffic"),
  time = c("08:00:00", "09:15:00", "10:30:00", "13:00:00", "15:45:00", "17:30:00")
)

# Convert time to POSIXct format
ddfw_filtered$time <- as.POSIXct(ddfw_filtered$time, format = "%H:%M:%S")

# Define weights for each subtype
subtype_weights <- c("Accident" = 3, "Road closure" = 2, "Heavy traffic" = 1)

# Filter data by time between 6am and 6pm
ddfw_filtered <- subset(ddfw_filtered, format(time, "%H") >= "06" & format(time, "%H") < "18")

# Calculate the severity score for each JAM alert
ddfw_filtered$score <- with(ddfw_filtered, subtype_weights[subtype] * table(subtype)[subtype])

# Sum the scores to obtain the total severity score
total_score <- sum(ddfw_filtered$score)

# Plot the severity scores by subtype and type
library(ggplot2)
ggplot(ddfw_filtered, aes(x = subtype, y = score, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "JAM Severity Scores by Subtype and Type", x = "Subtype", y = "Score")

```

## Waze aggregate over multiple events (20 points)
8.1 uuid for one major accident is "68b55df9-30f0-4a7e-9556-7f11f366549b"
```{r}

selected_accident <- dfw |>
  filter(type == "ACCIDENT", subtype == "ACCIDENT_MAJOR", uuid == "68b55df9-30f0-4a7e-9556-7f11f366549b")

# Longitude and latitude of accident location
accident_longitude <- selected_accident$longitude
accident_latitude <- selected_accident$latitude

# Defining the geographic box around the accident location with a 1000 meter radius
box_radius <- 1000

# Defining time range (one hour before and one hour after the accident)
accident_time <- selected_accident$ts_ct_rounded
time_range_start <- accident_time - lubridate::hours(1)
time_range_end <- accident_time + lubridate::hours(1)

# Filtering by time range and geographic box
sample_alerts <- dfw |>
  filter(
    ts_ct_rounded >= time_range_start & ts_ct_rounded <= time_range_end,
    geosphere::distVincentySphere(
      cbind(longitude, latitude),
      c(accident_longitude, accident_latitude)
    ) * 6371000 <= box_radius
  )

# Getting UUIDs of the sample alerts
sample_uuids <- sample_alerts$uuid

# Print the UUID of the selected accident and the sample UUIDs of alerts
print(sample_uuids)

# Filter the alert based on the time range and alert type
traffic_jam_alerts <- dfw |>
  filter(ts_ct_rounded >= time_range_start, 
         ts_ct_rounded <= time_range_end, type == "JAM")

# Create a five-minute interval variable
traffic_jam_alerts <- traffic_jam_alerts |>
  mutate(interval = cut(ts_ct_rounded, breaks = "5 min"))

# Counting the number of traffic jam alerts in each interval
alert_counts <- traffic_jam_alerts |>
  count(interval)

# Plotting the number of traffic jam alerts over time
ggplot(alert_counts, aes(x = interval, y = n)) +
  geom_point() +
  geom_line() +
  labs(x = "Time Interval", y = "Number of Traffic Jam Alerts") +
  ggtitle("Number of Traffic Jam Alerts Over Time") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```
8.2 

```{r}
install.packages("geosphere")
library(geosphere)

get_alerts_within_buffer <- function(dfw, input_uuid, buffer_distance) {
  selected_accident <- dfw %>% filter(uuid == input_uuid)
  
  accident_longitude <- selected_accident$longitude
  accident_latitude <- selected_accident$latitude
  accident_location <- c(accident_longitude[1], accident_latitude[1])  
  
  time_range_start <- selected_accident$ts_ct_rounded - hours(1)
  time_range_end <- selected_accident$ts_ct_rounded + hours(1)
  
  filtered_data <- dfw %>%
    filter(ts_ct_rounded >= time_range_start, ts_ct_rounded <= time_range_end)|>
    filter(geosphere::distVincentySphere(cbind(longitude, latitude), 
                                         accident_location) <=     
  buffer_distance) %>%
    mutate(interval = cut(ts_ct_rounded, breaks = "5 min")) |>
    count(interval)
  
  return(filtered_data)
}

uuid <- "68b55df9-30f0-4a7e-9556-7f11f366549b"
buffer_distance <- 10000  
result <- get_alerts_within_buffer(dfw, uuid, buffer_distance)

print(result)

```

