---
title: "applied_PS4"
author: "Matthew S."
date: "13th May 2023"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}

install.packages("readr")
library(readr)
dfw <- read_rds("/Users/mattida6/Desktop/GitHub/waze_data.rds")

```

# Front matter

This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: MS

Upload your collaborators here: https://forms.gle/3mH1ofZZj9ZCgBHb8

Late coins used this pset: 0. Late coins left: 4. <!--You may use one for a given assignment.-->

# R for Data Science Exercises

<!--Note: Please do not restate questions in your submissions, directly write your answer--->

```{r}

```

## 1 Prelim questions
## 2 git practice

# 1

The merge conflict occurred because both Partner A and Partner B made changes 
to the same line of the file in their respective branches, which conflicted 
with each other. As a result, when Partner A tried to merge Partner B's 
changes into their branch, there was a conflict because the same line was 
changed in two different ways. This requires manual intervention to resolve 
the conflict by editing the file to incorporate both changes.

## 3 Obtaining data from cloud using SQL
## 4 Data exploration in JSON
## 5 Data cleaning 
## 6 Waze vision zero
## 7 Waze single event
## 8 Waze aggregate over multiple events 
##