---
title: "applied_PS4"
author: "Matthew S."
date: "13th May 2023"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
rm(list=ls())

install.packages("bigrquery")
library(bigrquery)
library(tidyverse)

billing_and_proj_id <- "ganong-teaching-343718"

sql <- "SELECT
  *
FROM `waze-public-dataset.partner_UniversityofChicago.view_alerts_clustered`
WHERE city = 'Chicago, IL' AND date(ts) = '2023-02-01'"

tbl_waze <- bq_project_query(billing_and_proj_id, sql)

system.time(all_alerts <- bq_table_download(tbl_waze, n_max = 10))


```

<!-- .Rmd files use  markdown, a text mark up language, to provide formating.-->

<!--Text include within these strange arrows are comments and will not show up when you knit-->

# Front matter

This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: \<\<\_\>\>

Upload your collaborators here: https://forms.gle/3mH1ofZZj9ZCgBHb8

Late coins used this pset: 0. Late coins left: 4. <!--You may use one for a given assignment.-->

# R for Data Science Exercises

<!--Note: Please do not restate questions in your submissions, directly write your answer--->

## First Section

<!--(Notice the use of two `##`)-->

1.  YOUR TEXT ANSWER GOES HERE (Please delete)

-   The dash symbol creates unordered list

-   Another example

    1.  numbering creates ordered list
    2.  Another example

    -   like This
    -   Like this as well

**Delete all the text in `1.` before you write your answer**

```{r}
# YOUR CODE GOES HERE  (Please delete)
ggplot(mpg, aes(manufacturer)) + geom_bar()
```


## Second Section

1.  YOUR TEXT ANSWER GOES HERE (Please delete)

```{r}
# YOUR CODE GOES HERE  (Please delete)
ggplot(mtcars, aes(factor(cyl), mpg)) + geom_violin(aes(fill=mpg))
```

